---
title: "Resignation"
author: "Yph Lelkes"
date: "2022-12-08"
output: tint::tintHtml
latexfonts: 
  - package: newtxmath
    options: 
      - cmintegrals
      - cmbraces
  - package: ebgaramond-maths
  - package: nimbusmononarrow

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,warning = F,message = F)
library(tidyverse)
library(freqtables)
wrapper <- function(x, ...) 
{
  paste(strwrap(x, ...), collapse = "\n")
}

```

Summary: While lots of people are worried about the future of democracy, few think its end is inevitable (roughly 17 percent of the population).

1. Resignation is similar across gender, race, but the more educated are less resigned. 
2. Republicans are more resigned than democrats, pure independents are the most resigned
3. Resignation is related to a bunch of political attitudes. 
  * slightly less trusting, generally
  * Those who are resigned are less likely to think voting is important
  * Don't think they'll get fair treatment from the government
  * not particularly proud to be american
  * don't think the government will come to their aid
  * think the government is corrupt
4. Very little difference in perceptions of the other side's support for democratic norms
5. But they are more likely to support undemocratic behavior
6. And far more likely to support partisan violence
7. Resigned democrats, but not resigned republicans, are less affectively polarized. 


```{r}
load("resignation.RData")
questions <- read.csv("../../survey data/questions.csv")
resignation$res1 <- factor(resignation$res1,levels=rev(c("Strongly disagree","Somewhat disagree","Neither agree nor disagree","Somewhat agree","Strongly agree")))
resignation$res2 <- factor(resignation$res2,levels=rev(c("Strongly disagree","Somewhat disagree","Neither agree nor disagree","Somewhat agree","Strongly agree")))
```

```{r}

resignation %>% count(res1,wt =weight ) %>% mutate(p=n/sum(n)) %>% ggplot(aes(x=res1,y = p,label = scales::percent(p))) + 
  geom_col() + 
  scale_y_continuous(labels=scales::percent) +
  ylab("relative frequencies") + theme_bw(base_size = 15)+ggtitle("America is heading toward the end of democracy, where free and fair elections will no longer occur.") +xlab("")+theme(plot.title = element_text(size = 10))+geom_text(position = position_dodge(width = .9),    # move to center of bars
              vjust = -0.5,    # nudge above top of bar
              size = 3) + 
    scale_y_continuous(labels = scales::percent)+theme(axis.text.x = element_text(angle = 90))+theme(plot.title = element_text(size = 10))


```

```{r}
resignation %>% count(res2,wt =weight ) %>% mutate(p=n/sum(n)) %>% drop_na() %>% ggplot(aes(x=res2,y = p,label = scales::percent(p))) + 
  geom_col() + 
  scale_y_continuous(labels=scales::percent) +
  ylab("relative frequencies") + theme_bw(base_size = 15)+ggtitle("Very little can be done to stop American democracy from ending in the next decade.") +xlab("")+geom_text(position = position_dodge(width = .9),    # move to center of bars
              vjust = -0.5,    # nudge above top of bar
              size = 3) + 
    scale_y_continuous(labels = scales::percent)+    scale_y_continuous(labels = scales::percent)+theme(axis.text.x = element_text(angle = 90))+theme(plot.title = element_text(size = 10))



```

## Resignation: Agreeing with both comments (roughly 17 percent of the population)

```{r}
resignation$resigned <- ifelse(as.numeric(resignation$res1)<3 & as.numeric(resignation$res2)<3,"Resigned","Not Resigned") 
resignation %>% count(resigned,wt =weight ) %>% mutate(p=n/sum(n)) %>% drop_na() %>% ggplot(aes(x=resigned,y = p,label = scales::percent(p))) + 
  geom_col() + 
  scale_y_continuous(labels=scales::percent) +
  ylab("relative frequencies") + theme_bw(base_size = 15)+ggtitle("Percentage of Americans who are resigned") +xlab("")+    geom_text(position = position_dodge(width = .9),    # move to center of bars               vjust = -0.5,    # nudge above top of bar               size = 3) +      scale_y_continuous(labels = scales::percent)+    geom_text(position = position_dodge(width = .9),    # move to center of bars
              vjust = -0.5,    # nudge above top of bar
              size = 3) + 
    scale_y_continuous(labels = scales::percent)


```


## Who is resigned?

```{r}
resignation <- resignation %>% mutate(race = case_when(race=='White'~'White',
                                        race=='Black'~'Black',
                                        race=='Hispanic'~'Hispanic',
                                        TRUE~'Other'))

resignation %>% group_by(race) %>% freq_table(race,resigned) %>%  ggplot(aes(x=row_cat,y = percent_row))+geom_col(position = "dodge") + geom_errorbar(aes(ymin=lcl_row,ymax=ucl_row),position=position_dodge(width=0.9))+ggtitle("Resignation by Race")+facet_wrap(~col_cat)
```

```{r}
resignation %>% group_by(gender) %>% freq_table(gender,resigned) %>%  ggplot(aes(x=row_cat,y = percent_row))+geom_col(position = "dodge") + geom_errorbar(aes(ymin=lcl_row,ymax=ucl_row),position=position_dodge(width=0.9))+ggtitle("Resignation by Gender")+facet_wrap(~col_cat)
```

### Percent resigned by education
```{r}
resignation <- resignation %>% mutate(educ = case_when(educ %in% c("No HS","High school graduate")~'HS or Less',
                                        educ %in% c("Some college","2-year")~'Some College',
                                        educ %in% c("4-year","Post-grad")~'College+')) %>% mutate(educ = factor(educ,levels=c('HS or Less','Some College','College+')))

resignation %>% group_by(educ) %>% freq_table(educ,resigned) %>%  ggplot(aes(x=row_cat,y = percent_row))+geom_col(position = "dodge") + geom_errorbar(aes(ymin=lcl_row,ymax=ucl_row),position=position_dodge(width=0.9))+ggtitle("Resignation by Education")+facet_wrap(~col_cat)

```
### Partisanship
```{r}
resignation <- resignation %>% mutate(pid=case_when(grepl("Democrat",pid7) ~'Democrat',
                                     grepl("Republican",pid7) ~'Republican',
                                     TRUE~'Independent'))
                                     
resignation %>% group_by(pid) %>% freq_table(pid,resigned) %>%  ggplot(aes(x=row_cat,y = percent_row))+geom_col(position = "dodge") + geom_errorbar(aes(ymin=lcl_row,ymax=ucl_row),position=position_dodge(width=0.9))+ggtitle("Resignation by PID")+facet_wrap(~col_cat)
```

### Partisanship
```{r}
resignation <- resignation %>% mutate(presvote20post_1 = case_when(presvote20post=="Donald Trump"~'Trump',
                                                  presvote20post=="Joe Biden" ~ 'Biden',
                                                  presvote20post=='Did not vote for President'~'Did not vote',
                                                  TRUE~'Other'))
resignation <- resignation %>% mutate(pid=case_when(grepl("Democrat",pid7) ~'Democrat',
                                     grepl("Republican",pid7) ~'Republican',
                                     TRUE~'Independent'))
                                     
resignation %>% group_by(presvote20post_1) %>% freq_table(presvote20post_1,resigned) %>%  ggplot(aes(x=row_cat,y = percent_row))+geom_col(position = "dodge") + geom_errorbar(aes(ymin=lcl_row,ymax=ucl_row),position=position_dodge(width=0.9))+ggtitle("Resignation by PID")+facet_wrap(~col_cat) + xlab("")

## Do you think your wallet (or your valuables) would be returned to you if it were found by a stranger?
```

```{r}
resignation$general_trust[resignation$general_trust==""]=NA
resignation %>% drop_na(general_trust) %>% group_by(general_trust) %>% freq_table(general_trust,resigned) %>%  ggplot(aes(x=row_cat,y = percent_row))+geom_col(position = "dodge") + geom_errorbar(aes(ymin=lcl_row,ymax=ucl_row),position=position_dodge(width=0.9))+ggtitle(wrapper("Do you think your wallet (or your valuables) would be returned to you if it were found by a stranger?",40))+facet_wrap(~col_cat) + xlab("")
#  geom_col(position = "dodge2") + 
#  scale_y_continuous(labels=scales::percent) +
#  ylab("relative frequencies") + theme_bw(base_size = 15)+ggtitle("Percentage of Americans who are resigned") +xlab("") 
```

## 

```{r}
resignation$institutional_corruption <- factor(resignation$institutional_corruption,c("Extremely likely to refuse","Likely to refuse","Equally likely to refuse or accept","Likely to accept","Extremely likely to accept"))

resignation %>% drop_na(general_trust) %>% group_by(general_trust) %>% freq_table(institutional_corruption,resigned) %>%  ggplot(aes(x=row_cat,y = percent_row))+geom_col(position = "dodge") + geom_errorbar(aes(ymin=lcl_row,ymax=ucl_row),position=position_dodge(width=0.9))+ggtitle(wrapper("If you were to complain about the poor quality of a public service, how likely or unlikely is it that the problem would be easily resolved?",width=40))+facet_wrap(~col_cat) + xlab("")
#  geom_col(position = "dodge2") + 
```


```{r}
resignation$vote_importance <- factor(resignation$vote_importance,c("Very unimportant","Unimportant","Neither important nor unimportant","Important","Very important"))
resignation %>% drop_na(vote_importance) %>% group_by(vote_importance) %>% freq_table(vote_importance,resigned) %>%  ggplot(aes(x=fct_relevel(row_cat,c("Very unimportant","Unimportant","Neither important nor unimportant","Important","Very important")),y = percent_row))+geom_col(position = "dodge") + geom_errorbar(aes(ymin=lcl_row,ymax=ucl_row),position=position_dodge(width=0.9))+ggtitle(wrapper("How important or unimportant is it to vote in every election?",40))+facet_wrap(~col_cat) + xlab("")

```


## Percent resigned by pride in being American 
```{r}
resignation$pride <- factor(resignation$pride,c("Not at all proud","Only a little proud","Moderately proud","Very proud","Extremely proud"))

resignation %>% drop_na(pride) %>% freq_table(pride,resigned) %>%  ggplot(aes(x=fct_relevel(row_cat,levels(resignation$pride)),y = percent_row))+geom_col(position = "dodge") + geom_errorbar(aes(ymin=lcl_row,ymax=ucl_row),position=position_dodge(width=0.9))+ggtitle(wrapper("How important or unimportant is it to vote in every election?",40))+facet_wrap(~col_cat) + xlab("")


```

## Do you agree or disagree that you can expect fair treatment from government authorities?

```{r}
resignation$fair_treatment <- factor(resignation$fair_treatment,levels=rev(c("Strongly disagree","Disagree","Neither agree nor disagree","Agree","Strongly agree")))
resignation %>% drop_na(fair_treatment) %>% freq_table(fair_treatment,resigned) %>%  ggplot(aes(x=fct_relevel(row_cat,levels(resignation$fair_treatment)),y = percent_row))+geom_col(position = "dodge") + geom_errorbar(aes(ymin=lcl_row,ymax=ucl_row),position=position_dodge(width=0.9))+ggtitle(wrapper("How important or unimportant is it to vote in every election?",40))+facet_wrap(~col_cat) + xlab("")

#  geom_col(position = "dodge2") + 
#  scale_y_continuous(labels=scales::percent) +
#  ylab("relative frequencies") + theme_bw(base_size = 15)+ggtitle("Percentage of Americans who are resigned") +xlab("") 
```




### Resigned people are far more likely to support undemocratic behavior

```{r}
resignation %>% select(resigned,starts_with("norm")) %>% select(!ends_with("perception")) %>% pivot_longer(!resigned) %>% left_join(questions,c("name"="label")) %>% mutate(value1=case_when(
  value %in% c("Agree","Strongly Agree") ~ 1,
  value %in% c("Strongly disagree","Disagree","Neither agree nor disagree") ~ 0)) %>% drop_na() %>% group_by(short_label,resigned) %>% summarise(m=mean(value1),n=n())


#  geom_col(position = "dodge2") + 
#  scale_y_continuous(labels=scales::percent) +
#  ylab("relative frequencies") + theme_bw(base_size = 15)+ggtitle("Percentage of Americans who are resigned") +xlab("") 
```

## Percent resigned by support for violence
```{r}
resignation$violence2[resignation$violence2==""]=0
resignation$violence2 <- ifelse(resignation$violence2 %in% c("Strongly support","Support"),1,0)
resignation$violence3[resignation$violence3==""]=0
resignation$violence3 <- ifelse(resignation$violence3 %in% c("Strongly support","Support"),1,0)
resignation$violence4[resignation$violence4==""]=0
resignation$violence4 <- ifelse(resignation$violence4 %in% c("Strongly support","Support"),1,0)
resignation$violence5[resignation$violence5==""]=0
resignation$violence5 <- ifelse(resignation$violence5 %in% c("Strongly support","Support"),1,0)
resignation$violence6[resignation$violence6==""]=0
resignation$violence6 <- ifelse(resignation$violence6 %in% c("Strongly support","Support"),1,0)

resignation %>% select(resigned,starts_with("violence")) %>% select(!ends_with("1")) %>% pivot_longer(!resigned) %>% left_join(questions,c("name"="label")) %>% group_by(short_label,resigned) %>% summarise(m=mean(value),n=n())



```


```{r,eval=F}
resignation <- resignation %>% mutate(pid=case_when(grepl("Democrat",pid7) ~'Democrat',
                                     grepl("Republican",pid7) ~'Republican',
                                     TRUE~NA_character_))

resignation$inparty[which(resignation$pid == "Democrat")] <- resignation$democrat_therm_1[which(resignation$pid == "Democrat")]
resignation$inparty[which(resignation$pid == "Republican")] <- resignation$republican_therm_1[which(resignation$pid == "Republican")]
resignation$outparty[which(resignation$pid == "Republican")] <- resignation$democrat_therm_1[which(resignation$pid == "Republican")]
resignation$outparty[which(resignation$pid == "Democrat")] <- resignation$republican_therm_1[which(resignation$pid == "Democrat")]
resignation$affpol <- resignation$inparty - resignation$outparty


resignation %>% group_by(resigned,pid) %>% summarise(mean=mean(affpol,na.rm=T ),n=n()) %>% drop_na()
```
s